### 2022-07-25 doc/VerticalArrangements.tex
Reran figures using Dirichelt parameter of 1. Incorporated new numbers into paper. Minor other changes to paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title Page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Beyond ``Horizontal'' and ``Vertical'':\\ The Welfare Effects of Complex Integration\footnote{The analysis and conclusions set forth are those of the authors and do not indicate concurrence by other members of the Board research staff or by the Federal Reserve Board of Governors. Furthermore, the views expressed here should not be purported to reflect those of the Federal Trade Commission or the U.S. Department of Justice.}}

@@

The median average wholesale pre-merger price is \$5.80, and the median average retail pre-merger price is \$15.  Recall that the market size is set to 1, meaning that average price is equal to total pre-merger expenditures.  Pre-merger HHIs range from 2,911 at the $25^{th}$ percentile to 4,980 at the $75^{th}$, with a median of 3,747. HHIs for vertical mergers increase by 1,328 points at the median, resulting in a median post-merger HHI equal to 6,045.\footnote{We compute the post-merger HHI for vertical mergers by treating the merged firms' market share as the sum of all the shares of downstream products that either use the upstream partner's input or are sold by the downstream partner.  }  HHIs for downstream mergers increase by 2,126 points at the median, resulting in a median post-merger HHI equal to 5,149.  HHIs for upstream mergers increase by 2,069 points at the median, resulting in a median post-merger HHI equal to 4,967. HHIs for integrated mergers increase by 2,801 points at the median, resulting in a median post-merger HHI equal to 7,135.

@@

Turning to wholesaler surplus in the third panel of Figure \ref{fig:surplussum}, the effects seen there are largely reversed from those for retailers: wholesaler surplus increases in about 64\% of all upstream mergers, 11\% of vertical mergers, and about 16\% of downstream  and integrated mergers. Here, wholesaler surplus under integrated, downstream and upstream mergers each stochastically dominate the surplus from vertical mergers, but not one another. The rank ordering  of median harm, however, is the reverse of that from

@@

We begin with vertical mergers, as they are our simplest case, because there is never any pre-existing integration at either merging firm.  Absent incumbent integration by rivals (denoted by 0 integrated firms in Figure \ref{fig:CVvertincumbBW_consumer}), about half of vertical mergers benefit consumers and about 31\% of vertical mergers increase total surplus.  The addition of a single rival integrated firm (1 integrated firm in the figure) increases median consumer harm from 0.2\% to to 2\% while widening the distribution of outcomes slightly. By contrast, additional integrating rivals has little effect on median total harm. Adding more integrated rivals tends to increase the proportion of mergers that benefit consumers, while also increasing the proportion of mergers that reduce total surplus. Thus, the presence of rival integrated firms can limit harm, particularly consumer harm, from vertical mergers.  Integrated rivals have sources of bargaining leverage that mirror the additional leverage a vertical merger brings to the combining firms.  It appears that these rivals can therefore somewhat limit the extent of negative impacts that would otherwise occur through RRC.

Moving to the second panel from the left in Figure \ref{fig:CVvertincumbBW_consumer}, upstream mergers absent incumbent vertical integration (0 integrated firms in the figure) are never beneficial in the range we study, with median consumer harm at about 6\% of pre-merger revenues and median total harm at about 3\% of pre-merger revenues. This result is to be expected, as upstream mergers without additional integration do not have the potential countervailing effects for welfare that have been discussed in the context of downstream or vertical mergers.  Once one of the merging firms is allowed to be integrated (1 integrated firm in the figure) beneficial mergers appear, and median harm changes modestly.  Consumer welfare increases in almost 30\% of mergers and total welfare increases in about 10\% of mergers, but median consumer harm and total harm are roughly steady at approximately 4\% pre-merger revenues.  The additional vertical integration creates some opportunities for EDM to enhance welfare, but this must be balanced against the potential for RRC.  Just as we saw with downstream mergers, here we find that harm dominates in most simulations.  Adding in integration by rival firms (2-4 integrated firms in the figure) narrows the inter-quartile range and trims the lower whisker-- meaning that relatively more harmful mergers are eliminated than beneficial ones-- while only slightly changing median harm.  Therefore, the presence of third-party incumbent integrated firms, as in the vertical and downstream cases, can limit the likelihood of observing extremely harmful mergers.

Next we examine the welfare impacts of downstream mergers. As with upstream mergers,  downstream mergers absent incumbent vertical integration (0 integrated firms in the figure) are never beneficial in the $5^{th}$ to $95^{th}$ percentile range, with median consumer harm from downstream mergers equal to about 9\% of pre-merger revenues and total harm equal to about 5\% of pre-merger revenues. \citeN{ST2021} found the same result for mergers in downstream logit Bertrand markets.  Allowing one of the merging firms to be vertically integrated when all other market participants are unintegrated (1 integrated firm in the figure) increases the range of outcomes while also leading to more median harm: consumer welfare increases in about 26\% of mergers, and total welfare increases in about 6\%, but median consumer harm grows to about 11\% of pre-merger revenues, and median total harm grows to about 7\%. Once one of the merging firms is already integrated, the merger now has potential EDM and RRC effects, which raises the possibility of both benefits and harms to welfare.  On net, we find that the harms dominate in most instances.  Adding integrated third party rivals (2-4 integrated firms in the figure) somewhat decreases the range of outcomes and moves the distribution towards less harm.  With 3 integrated third parties (4 integrated firms in the figure), the merger benefits consumers in about 34\% of markets and shrinks median consumer harm to about 5\% of pre-merger revenues. The inter-quartile range of total harm narrows, and median total harm falls to about 4\% of pre-merger revenues. As we saw with vertical mergers, it appears that the presence of third party integrated rivals can limit harm, though here the effect is a bit stronger.

We finish by examining integrated mergers in the right-most panel of Figure \ref{fig:CVvertincumbBW_consumer}.  Absent the presence of rival incumbent integrated firms (2 integrated firms in the figure), mergers between two integrated firms benefits consumers in about 8\% of simulated markets and are net beneficial in 15\% of simulated markets. Thus, we find that mergers between two firms that are already integrated are harmful in the vast majority of cases.  While adding one additional rival incumbent integrated firm (3 integrated firms in the figure) has little effect on the distribution of either consumer or total welfare, adding a second integrated rival (4 integrated firms in the figure) eliminates the most harmful integrated mergers by shortening the lower whisker. Adding further integrated incumbents has modest incremental effects on the distribution of harm.

To summarize, our analysis of Figure \ref{fig:CVvertincumbBW_consumer} yields some overall conclusions.  First, we find that mergers where one or both of the merging firms are already integrated (and no rivals are integrated) are only beneficial in a small proportion of instances.  Second, the presence of integrated rivals can decrease the likelihood of observing the most harmful mergers, though the strength of that effect can vary widely across types of mergers.

@@

Figure \ref{fig:CVbargincumbent_consumer}  depicts box and whisker plots summarizing the consumer (top panels) and total (bottom panels) welfare effects for downstream, upstream, and vertical mergers as the relative bargaining power parameter goes from 9 (wholesalers have the advantage) to 1/9 (retailers have the advantage). Also depicted for each bargaining power parameter are three sets of box and whisker plots that correspond to the number of incumbent vertically integrated firms included in the simulated markets: 0 (light blue), 1 (darker blue), and 4 (darkest blue) firms.  As in Figure \ref{fig:CVvertincumbBW_consumer}, plots when the number of integrated firms is 0 assume that pre-merger, no firms in the market are vertically integrated; these are comparable to those in \citeN{ST2021}. For vertical mergers, the plots when the number of incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when one third party rival is integrated. By contrast, for upstream and downstream mergers, the plots when the number of incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated firm when all third parties are unintegrated. In turn, plots with 4 integrated firms increase the number of rival incumbent integrated firms in vertical mergers to 4 and in upstream and downstream mergers to 3.  Figure \ref{fig:CVbargincumbent_consumer} provides analogous results for integrated mergers.  In integrated mergers, when the number of  incumbent integrated firms equals 2, this depicts a merger between two integrated firms with all rivals are unintegrated. Plots with either 4 or 6 integrated firms depict mergers between integrated incumbents when there are either 2 or 4 rival integrated firms in the market, respectively.

We begin by examining vertical mergers.  We find that the presence of integrated incumbents has little impact on the relationship between relative bargaining power and welfare. Consistent with the patterns seen in \citeN{ST2021}, mergers tend to benefit welfare when wholesalers have relatively more power, and harm welfare when retailers have relatively more power.  Larger wholesaler bargaining power offers more possibilities for EDM, as pre-merger input prices are likely to be high.  This channel becomes less relevant as retailers gain more power.  The impact on consumers is roughly zero when wholesalers and retailers have equal power.  Furthermore, we find that consumer harm when there are fewer incumbent integrated firms first-order stochastically dominates harm when there are more incumbent integrated firms, although the magnitude of these differences is small. Across all three integrated incumbent scenarios, median consumer harm moves from about -24\% of pre-merger expenditures when wholesalers have the advantage to 7\% when retailers have the advantage.  Similarly, median total harm goes from from about -14\% of pre-merger expenditures when wholesales have the advantage to 2.9\% when retailers have the advantage.

We turn next to upstream mergers in Figure \ref{fig:CVbargincumbent_consumer}.  Without incumbent integrated firms in the market (0 integrated firms), the relationship between harm and bargaining power is reversed relative to downstream and vertical mergers: harm is greater when wholesalers have relatively more bargaining power.  Median consumer harm moves from about 28\% of pre-merger expenditures when wholesalers are more powerful to 2\% of pre-merger expenditures when retailers are more powerful. Likewise, median total harm goes from approximately 12\% to less than 1\% of pre-merger expenditures. By contrast, for mergers between an integrated and an unintegrated upstream supplier (denoted by 1 integrated firm in the figure), the relationship between harm and bargaining power is more in line with that for downstream and vertical mergers.  This is intuitive, as now all three types of mergers create some form of vertical integration.  Median consumer harm goes from roughly -23\% of pre-merger expenditures when wholesalers have relatively more bargaining power to almost 8\% of pre-merger expenditures when retailers have relatively more bargaining power. Median total harm increases from about -20\% to more than 3\% of pre-merger expenditures. Adding incumbent integrated rivals (see 4 integrated firms in the figure) preserves the relationship between bargaining power and harm, though the spread of outcomes narrows.

For downstream mergers, we continue to find that higher retailer relative bargaining power leads to more negative impacts on welfare.  Absent any incumbent integration (0 integrated firms), all the outcomes shown in Figure \ref{fig:CVbargincumbent_consumer} feature a decrease in welfare, which mirrors our finding in Figure \ref{fig:CVvertincumbBW_consumer}.  Median consumer harm goes from about 1\% of pre-merger expenditures when wholesalers have relatively more bargaining power to 18\% of pre-merger expenditures when retailers have relatively more bargaining power. Likewise, median total harm moves from about 2\% of pre-merger expenditures to about 8\%.  In situations where retailers already have more bargaining power, retailers are likely to have extracted significant surplus from wholesalers prior to the merger, which limits any potential benefits from increased bargaining leverage.  Thus, there is little to counteract harms from decreased downstream competition.

Once we allow one of the merging firms to be integrated (denoted by 1 integrated firm in the figure), some beneficial mergers appear, particularly when wholesalers have relatively more bargaining power.  The box and whisker plots rotate clockwise around a pivot at equal bargaining power, strengthening the relationship between bargaining power and extent of harm.  Median consumer harm goes from approximately -33\% of pre-merger revenues when wholesalers have the advantage to 24\% of pre-merger revenues when retailers have the advantage.  Median total harm goes from about 2\% to almost 12\% of pre-merger revenues.  The existence of pre-merger integration at one of the merging firms creates an opportunity for EDM, the benefits from which are likely to be largest when wholesalers have high bargaining power and therefore charge high pre-merger input prices.  Compared to vertical mergers, where impacts to consumers were roughly neutral at 2/3 bargaining power, here neutrality happens between  3/2 and 7/3, when wholesalers have a bit more than 1.5 times the power of retailers.  It appears that wholesalers must be relatively more powerful (and thus the likely gains from EDM larger) in order to generate net consumer benefits, compared to in a vertical merger.  This is intuitive, as here the downstream merger also causes an additional lessening in horizontal competition.  Once we add rival integrated firms (see 4 integrated firms in the figure) the box and whisker plots rotate counter-clockwise, weakening but not erasing the relationship between bargaining power and harm. The range of outcomes also shrinks.  Starting from when wholesalers have relatively more bargaining power and moving right, median consumer harm goes from approximately -35\% of pre-merger revenues to 13\% of pre-merger revenues, while median total harm goes from almost 3\% to 5\% of pre-merger revenues.

Figure \ref{fig:CVbargincumbent_consumer} displays results for integrated mergers.  In terms of bargaining power and resulting harm, integrated mergers are perhaps most similar to vertical mergers. Like vertical mergers, there is a strong monotonic relationship between bargaining power and harm. Median consumer harm moves from less than -2\% of pre-merger expenditures when wholesalers have relatively more bargaining power to about 26\% of pre-merger expenditures when retailers have relatively more bargaining power. Likewise, median total harm moves from about -20\% when wholesalers have relatively more bargaining power to 15\% of pre-merger expenditures when retailers have relatively more bargaining power.  A second similarity between vertical and integrated mergers is that the presence of rival integrated incumbents (see 4 or 6 integrated firms in the figure) has little impact on the relationship between bargaining power and harm. However, a key difference between vertical and integrated mergers is that while vertical mergers are often beneficial when wholesalers have relatively more bargaining power, integrated mergers are often harmful unless wholesaler relative bargaining power is roughly 4 or more. The crossing at zero harm is shifted to the left.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{output/CVbargincumbent_consumer.png}
\caption{The figure displays box and whisker plots summarizing the extent to which bertical, upstream, downstream and integrated mergers affect consumer surplus as the bargaining power of wholesalers relative to retailers changes. The different colored boxes display how outcomes change as the number of vertically  integrated firms increases. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median.}
\label{fig:CVbargincumbent_consumer}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{output/CVbargincumbent_total.png}
\caption{The figure displays box and whisker plots summarizing the extent to which bertical, upstream, downstream and integrated mergers affect total surplus as the bargaining power of wholesalers relative to retailers changes. The different colored boxes display how outcomes change as the number of vertically  integrated firms increases. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median.}
\label{fig:CVbargincumbent_total}
\end{sidewaysfigure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title Page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Beyond ``Horizontal'' and ``Vertical'':\\ The Welfare Effects of Complex Integration\footnote{The analysis and conclusions set forth are those of the authors and do not indicate concurrence by other members of the Board research staff or by the Federal Reserve Board of Governors. Furthermore, the views expressed here should not be purported to reflect those of the Federal Trade Commission or the U.S. Department of Justice. This article has benefited from conversations with Alison Oldale, Ted Rosenbaum, and Nathan Wilson.}}

@@

The median average wholesale pre-merger price is \$5.80, and the median average retail pre-merger price is \$15.  Recall that the market size is set to 1, meaning that average price is equal to total pre-merger expenditures.  Pre-merger HHIs range from 2,911 at the $25^{th}$ percentile to 4,981 at the $75^{th}$, with a median of 3,747. HHIs for vertical mergers increase by 1,328 points at the median, resulting in a median post-merger HHI equal to 6,048.\footnote{We compute the post-merger HHI for vertical mergers by treating the merged firms' market share as the sum of all the shares of downstream products that either use the upstream partner's input or are sold by the downstream partner.  }  HHIs for downstream mergers increase by 2,128 points at the median, resulting in a median post-merger HHI equal to 5,153.  HHIs for upstream mergers increase by 2,070 points at the median, resulting in a median post-merger HHI equal to 4,968. HHIs for integrated mergers increase by 2,803 points at the median, resulting in a median post-merger HHI equal to 6,998.

@@

Turning to wholesaler surplus in the third panel of Figure \ref{fig:surplussum}, the effects seen there are largely reversed from those for retailers: wholesaler surplus increases in about 64\% of all upstream mergers, 11\% of vertical mergers, and about 16\% of downstream  and integrated mergers. Here, wholesaler surplus under integrated, downstream and upstream mergers each stochastically dominate the surplus from vertical mergers, but not one another.

@@

We begin with vertical mergers, as they are our simplest case, because there is never any pre-existing integration at either merging firm.  Absent incumbent integration by rivals (denoted by 0 integrated firms in Figures \ref{fig:CVvertincumbBW_consumer} and \ref{fig:CVvertincumbBW_total}), about half of vertical mergers benefit consumers and about 31\% of vertical mergers increase total surplus.  The addition of a single rival integrated firm (1 integrated firm in the figure) decreases median consumer harm from about 0\% of pre-merger revenues to to 2\% while widening the distribution of outcomes slightly. By contrast, additional integrating rivals has little effect on median total harm. Adding more integrated rivals tends to increase the proportion of mergers that benefit consumers, while also increasing the proportion of mergers that reduce total surplus. Thus, the presence of rival integrated firms can limit harm, particularly consumer harm, from vertical mergers.  Integrated rivals have sources of bargaining leverage that mirror the additional leverage a vertical merger brings to the combining firms.  It appears that these rivals can therefore somewhat limit the extent of negative impacts that would otherwise occur through RRC.

Moving to the second panel from the left in Figures \ref{fig:CVvertincumbBW_consumer} and \ref{fig:CVvertincumbBW_total}, upstream mergers absent incumbent vertical integration (0 integrated firms in the figure) are never beneficial in the range we study, with median consumer harm at about 6\% of pre-merger revenues and median total harm at about 3\% of pre-merger revenues. This result is to be expected, as upstream mergers without additional integration do not have the potential countervailing effects for welfare that have been discussed in the context of downstream or vertical mergers.  Once one of the merging firms is allowed to be integrated (1 integrated firm in the figure) beneficial mergers appear, and median harm decreases modestly.  Consumer welfare increases in almost 30\% of mergers and total welfare increases in about 14\% of mergers, but median consumer harm and total harm are roughly steady at approximately 4\% of pre-merger revenues.  The additional vertical integration creates some opportunities for EDM to enhance welfare, but this must be balanced against the potential for RRC.  Just as we saw with downstream mergers, here we find that harm dominates in most simulations.  Adding in integration by rival firms (2-4 integrated firms in the figure) narrows the inter-quartile range and trims the lower whisker-- meaning that relatively more harmful mergers are eliminated than beneficial ones-- while only slightly changing median harm.  Therefore, the presence of third-party incumbent integrated firms, as in the vertical and downstream cases, can limit the likelihood of observing extremely harmful mergers.

Next we examine the welfare impacts of downstream mergers. As with upstream mergers,  downstream mergers absent incumbent vertical integration (0 integrated firms in the figure) are never beneficial in the $5^{th}$ to $95^{th}$ percentile range, with median consumer harm from downstream mergers equal to about 9\% of pre-merger revenues and total harm equal to about 6\% of pre-merger revenues. \citeN{ST2021} found the same result for mergers in downstream logit Bertrand markets.  Allowing one of the merging firms to be vertically integrated when all other market participants are unintegrated (1 integrated firm in the figure) increases the range of outcomes while also leading to more median harm: consumer welfare increases in about 20\% of mergers, and total welfare increases in about 2\%, but median consumer harm grows to about 12\% of pre-merger revenues, and median total harm grows to about 8\%. Once one of the merging firms is already integrated, the merger now has potential EDM and RRC effects, which raises the possibility of both benefits and harms to welfare.  On net, we find that the harms dominate in most instances.  Adding integrated third party rivals (2-4 integrated firms in the figure) somewhat decreases the range of outcomes and moves the distribution towards less harm.  With 3 integrated third parties (4 integrated firms in the figure), the merger benefits consumers in about 34\% of markets and shrinks median consumer harm to about 5\% of pre-merger revenues. The inter-quartile range of total harm narrows, and median total harm falls to about 4\% of pre-merger revenues. As we saw with vertical mergers, it appears that the presence of third party integrated rivals can limit harm, though here the effect is a bit stronger.

We finish by examining integrated mergers in the right-most panel of Figures \ref{fig:CVvertincumbBW_consumer} and\ref{fig:CVvertincumbBW_total}.  Absent the presence of rival incumbent integrated firms (0 integrated firms in the figures), mergers between two integrated firms benefits consumers in about 6\% of simulated markets and are net beneficial in about 17\% of simulated markets. Thus, we find that mergers between two firms that are already integrated are harmful in the vast majority of cases.  Adding one additional rival incumbent integrated firm (1 integrated firms in the figures) lengthens the upper whisker, increasing the percentage of beneficial mergers, but has little effect on either median consumer or total welfare. Likewise, adding a second integrated rival (2 integrated firms in the figures) lengthens the upper whisker while also eliminating the most harmful integrated mergers and slightly increasing median consumer and total harms. Adding further integrated incumbents has modest incremental effects on the distribution of harm.

To summarize, our analysis of Figures \ref{fig:CVvertincumbBW_consumer} and \ref{fig:CVvertincumbBW_total} yield some overall conclusions.  First, we find that mergers where one or both of the merging firms are already integrated (and no rivals are integrated) are only beneficial in a small proportion of instances.  Second, the presence of integrated rivals can decrease the likelihood of observing the most harmful mergers, though the strength of that effect can vary widely across types of mergers. Third, with the exception of downstream mergers,additional integrated firms often tend to increase the incidence of with the exception of integrated mergers, additional integrated firms do not substantively affects median consumer or total surplus, but can greatly effect outcome distribution skewness.

@@

Figures \ref{fig:CVbargincumbent_consumer}  and \ref{fig:CVbargincumbent_total}  depict box and whisker plots summarizing the consumer and total welfare effects for downstream, upstream, and vertical mergers as the relative bargaining power parameter goes from 9 (wholesalers have the advantage) to 1/9 (retailers have the advantage). Also depicted for each bargaining power parameter are three sets of box and whisker plots that correspond to the number of incumbent vertically integrated firms included in the simulated markets: 0 (light blue), 1 (darker blue), and 4 (darkest blue) firms.  As in Figures \ref{fig:CVvertincumbBW_consumer} and \ref{fig:CVvertincumbBW_total}, plots when the number of integrated firms is 0 assume that pre-merger, no firms in the market are vertically integrated; these are comparable to those in \citeN{ST2021}. For vertical mergers, the plots when the number of incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when one third party rival is integrated. By contrast, for upstream and downstream mergers, the plots when the number of incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated firm when all third parties are unintegrated. In turn, plots with 4 integrated firms increase the number of rival incumbent integrated firms in vertical mergers to 4 and in upstream and downstream mergers to 3.  Finally, in integrated mergers, when the number of  incumbent integrated firms equals 0, this depicts a merger between two integrated firms with all rivals are unintegrated. Plots with either 1 or 4 integrated firms depict mergers between integrated incumbents when there are either 1 or 4 rival integrated firms in the market, respectively.

We begin by examining vertical mergers.  We find that the presence of integrated incumbents has little impact on the relationship between relative bargaining power and welfare. Consistent with the patterns seen in \citeN{ST2021}, mergers tend to benefit welfare when wholesalers have relatively more power, and harm welfare when retailers have relatively more power.  Larger wholesaler bargaining power offers more possibilities for EDM, as pre-merger input prices are likely to be high.  This channel becomes less relevant as retailers gain more power.  The impact on consumers is roughly zero when wholesalers and retailers have equal power.  Furthermore, we find that consumer harm when there are 0 incumbent integrated firms first-order stochastically dominates harm when there are 4  incumbent integrated firms, with the difference in magnitudes decreasing as retailer bargaining power increases. Across all three integrated incumbent scenarios, median consumer harm moves from about -24\% of pre-merger expenditures when wholesalers have the advantage to 7\% when retailers have the advantage.  Similarly, median total harm goes from from about -14\% of pre-merger expenditures when wholesales have the advantage to 2.9\% when retailers have the advantage.

We turn next to upstream mergers in Figures \ref{fig:CVbargincumbent_consumer} and   \ref{fig:CVbargincumbent_total}.  Without incumbent integrated firms in the market (0 integrated firms), the relationship between harm and bargaining power is unlike that of downstream and vertical mergers: median harm first increases and then decreases as wholesalers have relatively more bargaining power, with the largest depicted median harm-- 12\% for consumers, 18\% on net-- occurring when relative bargaining power 7/3.  Increasing bargaining power beyond 7/3 evidently diminishes harm because pre-merger, wholesalers have already expropriated much of the retailers' surplus. By contrast, for mergers between an integrated and an unintegrated upstream supplier (denoted by 1 integrated firm in the figures), the relationship between harm and bargaining power is more in line with that for downstream and vertical mergers.  This is intuitive, as now all three types of mergers create some form of EDM that is to varying degrees passed through to consumers. Median consumer harm goes from roughly -21\% of pre-merger expenditures when wholesalers have relatively more bargaining power to almost 8\% of pre-merger expenditures when retailers have relatively more bargaining power. Median total harm increases from about -24\% to more than 3\% of pre-merger expenditures. Adding incumbent integrated rivals (see 4 integrated firms in the figure) preserves the relationship between bargaining power and harm, though the spread of outcomes narrows.

For downstream mergers, we continue to find that higher retailer relative bargaining power leads to more negative impacts on welfare.  Absent any incumbent integration (0 integrated firms), all the outcomes shown in Figures \ref{fig:CVbargincumbent_consumer} and \ref{fig:CVbargincumbent_total} feature a decrease in welfare, which mirrors our findings in Figures \ref{fig:CVvertincumbBW_consumer} and \ref{fig:CVvertincumbBW_total}.  Median consumer harm goes from about 1\% of pre-merger expenditures when wholesalers have relatively more bargaining power to 19\% of pre-merger expenditures when retailers have relatively more bargaining power. Likewise, median total harm moves from about 3\% of pre-merger expenditures to about 8\%.  In situations where retailers already have more bargaining power, retailers are likely to have extracted significant surplus from wholesalers prior to the merger, which limits any potential benefits from increased bargaining leverage.  Thus, there is little to counteract harms from decreased downstream competition.

Once we allow one of the merging firms to be integrated (denoted by 1 integrated firm in the figure), some beneficial mergers appear, particularly when wholesalers have relatively more bargaining power.  The box and whisker plots in Figure \ref{fig:CVbargincumbent_consumer} fold towards one another around a pivot at equal bargaining power, strengthening the relationship between bargaining power and extent of consumer harm.  Median consumer harm goes from approximately -28\% of pre-merger revenues when wholesalers have the advantage to 24\% of pre-merger revenues when retailers have the advantage.  Median total harm goes from almost 5\% to about 11\% of pre-merger revenues.  The existence of pre-merger integration at one of the merging firms creates an opportunity for EDM, the benefits from which are likely to be largest when wholesalers have high bargaining power and therefore charge high pre-merger input prices.  Compared to vertical mergers, where impacts to consumers were roughly neutral at 2/3 bargaining power, here neutrality happens between  3/2 and 7/3, when wholesalers have a bit more than 1.5 times the power of retailers.  It appears that wholesalers must be relatively more powerful (and thus the likely gains from EDM larger) in order to generate net consumer benefits, compared to in a vertical merger.  This is intuitive, as here the downstream merger also causes an additional lessening in horizontal competition.  Once we add rival integrated firms (see 4 integrated firms in the figure) the box and whisker plots rotate counter-clockwise, weakening but not erasing the relationship between bargaining power and harm. The range of outcomes also shrinks.  Starting from when wholesalers have relatively more bargaining power and moving right, median consumer harm goes from approximately -35\% of pre-merger revenues to 13\% of pre-merger revenues, while median total harm goes from almost 3\% to 5\% of pre-merger revenues.

The last column in Figures \ref{fig:CVbargincumbent_consumer} and   \ref{fig:CVbargincumbent_total} display results for integrated mergers.  In terms of bargaining power and resulting harm, integrated mergers are perhaps most similar to vertical mergers. Like vertical mergers, there is a strong monotonic relationship between bargaining power and harm. Median consumer harm moves from about 0\% of pre-merger expenditures when wholesalers have relatively more bargaining power to about 26\% of pre-merger expenditures when retailers have relatively more bargaining power. Likewise, median total harm moves from about -19\% when wholesalers have relatively more bargaining power to 15\% of pre-merger expenditures when retailers have relatively more bargaining power.  A second similarity between vertical and integrated mergers is that the presence of rival integrated incumbents (see 1 or 4 integrated firms in the figure) has modest impact on the relationship between bargaining power and harm. However, a key difference between vertical and integrated mergers is that while vertical mergers are often beneficial when wholesalers have relatively more bargaining power, integrated mergers are often harmful unless wholesaler relative bargaining power is roughly 4 or more. The crossing at zero harm is shifted to the left.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{output/CVbargincumbent_consumer.png}
\caption{The figure displays box and whisker plots summarizing the extent to which vertical, upstream, downstream and integrated mergers affect consumer surplus as the bargaining power of wholesalers relative to retailers changes. The different colored boxes display how outcomes change as the number of vertically  integrated firms increases. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median.}
\label{fig:CVbargincumbent_consumer}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{output/CVbargincumbent_total.png}
\caption{The figure displays box and whisker plots summarizing the extent to which vertical, upstream, downstream and integrated mergers affect total surplus as the bargaining power of wholesalers relative to retailers changes. The different colored boxes display how outcomes change as the number of vertically  integrated firms increases. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median.}
\label{fig:CVbargincumbent_total}
\end{sidewaysfigure}
